---
title: "Dimension reduction of categorical survey data using the R package `dhlvm`"
author: "Evan Munro"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`dhlvm` is an R package for estimating and analyzing the output of hierarchical latent variable models for both time series and cross-sectional survey data. The models are described in more detail in the following paper. This vignette uses the notation in the paper, and provides an overview of the R package functionality. 

For this tutorial, the `dhlvm` package includes a raw survey dataset `inflat`. It includes the responses of 2,000 individuals to 5 questions from the Bank of England Inflation Attitudes Survey. The responses are quarterly, from June XX to July XX. 

### Data cleaning 

For a dataset with $N$ survey respondents and $J$ questions, `dhlvm` expects two data inputs. The first is an $N x J$ matrix of survey responses, where each $X_{ij}$ is a numeric value from 1 to $L_j$ where $L_j$ is the total number of responses. This is a numeric score that corresponds to a categorical response. The function 

### Static model estimation

First, we ignore that the responses to the BoE inflation survey data was collected over time, and we group individuals by education level and estimate an LDA-S model. The key inputs are

### Dynamic model estimation 

Next, we recognize that a model that takes dynamics into account is better suited for survey data collected over time. To estimate a dynamic model, we group individuals by time instead of by XX. We then estimate an LDA-DS model. The inputs compared to the static model are the same, except that we provide an Inverse-Gamma prior for the variance of the random walk process in the lognormal distribution of $\pi_t$, rather than a Dirichlet prior for the multinomial distribution of $\pi_g$ in the static case. 



### Analyzing model posteriors 

The functions for analyzing the models are the same for both the dynamic and static models. The estimation procedure returns the full set of posterior draws for each of the parameters in the model. The first step is to calculate the posterior mean of all the parameters in the model: 

```{r} 

``` 

Then, functions are available for plotting the estimated $\pi$, the group/time-specific distribution over classes, either alone or against another data series. 

In addition, there are functions for plotting $\beta$, the class-specific distribution over responses for each question. 

For selection 
